---
title: "Inference in Shift-Share Designs"
author: "Michal Kolesár"
date: "`r Sys.Date()`"

bibliography: bartik_library.bib
output:
  pdf_document:
    toc: false
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Inference in Bartik Designs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, cache=FALSE}
library("knitr")
knitr::opts_knit$set(self.contained = FALSE)
knitr::opts_chunk$set(tidy = TRUE, collapse=TRUE, comment = "#>",
                      tidy.opts=list(blank=FALSE, width.cutoff=55))
```

The package `BartikSE` implements confidence intervals proposed by @akm18 for
inference in shift-share least squares and instrumental variables regressions,
in which the regressor of interest (or the instrument) has a shift-share
structure, as in @bartik91. A shift-share variable has the structure
$X_{i}=\sum_{s=1}^{S}w_{is}\mathcal{X}_{s}$, where $i$ indexes regions, $s$
indexes sectors, $\mathcal{X}_{s}$ are sectoral shifters (or shocks), and
$w_{is}$ are shares, such as initial share of region $i$'s employment in sector
$s$.

This vignette illustrates the use of the package using a dataset from @adh13
(ADH hereafter). The dataset is included in the package as the list `ADH`. The
first element of the list, `ADH$reg` is a data-frame with regional variables,
the second element, `ADH$sec` is a data-frame with sectoral variables (with
column `X` correspondong to the shifters), and `ADH$W` is a matrix of shares.
See `?ADH` for a description of the dataset.

We now replicate the first row of Table 6 in @akm18. First we load the package,
define the vector of controls, and define a vector of 3-digit SIC codes:
```{r}
library("BartikSE")
ctrls <- "t2 + l_shind_manuf_cbp + l_sh_popedu_c + l_sh_popfborn +
          l_sh_empl_f + l_sh_routine33 + l_task_outsource + division"
sic <- floor(ADH$sec$sic/10)
```

The first-stage regression:
```{r}
lmBartik(as.formula(paste("shock ~ ", ctrls)), W=ADH$W, Xs=ADH$sec$X,
               data=ADH$reg, weights=weights, region_cvar=statefip,
               sector_cvar=sic, method="all", residual_sector=TRUE)
```

The reduced-form and IV regressions:
```{r}
lmBartik(as.formula(paste("d_sh_empl ~", ctrls)), W=ADH$W, Xs=ADH$sec$X,
                  data=ADH$reg, region_cvar=statefip, weights=weights,
                  sector_cvar=sic, method="all", residual_sector=TRUE)
ivBartik(as.formula(paste("d_sh_empl ~", ctrls, "| shock")), W=ADH$W,
                  Xs=ADH$sec$X, data=ADH$reg, region_cvar=statefip,
                  weights=weights, sector_cvar=sic, method="all",
                  residual_sector=TRUE)
```

# References
